---
title: "Best Investment Strategy"
author: "Stephanie K Moore"
date: "3/3/2016"
output: html_document
---
##  Invest to Earn $5 Million in 20 Years   

                      My Scenario     
                      
Roll one number to apply to all risk categories - my thinking being that either it's a good year for all or it's not.     
I want to have at least 5 million in 20 years for retirement.     
I will start with $5000 due to my age.    
When I reach my goal of $5 million, I'll change my investment strategy to keep my $5 million safe.      


*This code built off of Warren Buffet Move Over. This time I'll try and step thru all combos of R Y and G.*   

```{r}
# Investment Portfolios, Numbers Cannot Be Changed #

R <- c(.05,.2,1,3,3,3)         # high risk, high return
Y <- c(.8,.9,1.1,1.1,1.2,1.4)  # med risk, med return
G <- c(.95,1,1,1,1,1.1)        # low risk, low return

# Initialize some counters etc.
goal <- 5000000
n <- 1000 # number of trials 
t <- 20 # number of years
yr <- numeric(t+1)  
# peryr[1] is start of 20 years; peryr[21] is end of 20th year
yr[1] <- 5000 # Init amount for each trial

tresults <- numeric(n) # store results of "n" trials

# Set up matrix to store results of Ratios use (cols 1-3), number of trials earning < $1 million (col 4), number of trials earning at least $5 million (col 5), year where $5 million was hit (if any) - col 6
rdata <- matrix(nrow = 1000, ncol = 6) 
```

```{r}
# Loops: Outer to Inner --> each ratio (k), for n trials (j), for 20 years (i)
for (k in 1:2) { 
  perR <- 0
  while (perR <= 1) { 
    perY <- 0
    while (perY <= min(1,perR)) {
      perG <- 0
      while (perG <= min(1,perR)) { 
          for (j in 1:n) {
            for (i in 1:t) {
              roll <- sample(1:6,1,replace=TRUE) # roll 1 number and apply to R, G, and Y
              yr[i+1] <- (R[roll]*perR + Y[roll]*perY + G[roll]*perG)*yr[i]
            }
            tresults[j] <- yr[i+1]
          }
          rdata[k,1] <- perR 
          rdata[k,2] <- perY 
          rdata[k,3] <- perG 
          rdata[k,4] <- sum(tresults < 1000000) 
          rdata[k,5] <- sum(tresults >= goal)
          rdata[k,6] <- t
          perG <- perG + 0.1
      }
      perY <- perY + 0.1
      }
    perR <- perR + 0.1
  }
}
rdata[1:100,1:4]
yr
tresults[1:10]
roll
perG
```

```{r}
sessionInfo()
```
