---
title: 'Shonkwiler Problem #13'
author: "Stephanie K Moore"
date: "`r Sys.Date()`"
output: html_document
---
## HIV Specificity and Sensitivity ##   

13. An early detection test for HIV is 98% accurate.  In a certain community, 3% of the population has HIV.  What is the probability that a random person with a positive test for HIV actually has the disease?  Simulate this by selecting a person at random from the population and deciding whether or not the person has HIV.  Then administer the test and determine whether is was positive.  Keep a counter for the times that the test is positive and the times the test is positive AND the person has HIV.   

```{r}
sens <- .98
spec <- .98
ispos <- .03
isneg <- .97
### Theoretical Probabilities ###
# True Positive
truepos <- sens*ispos
# True Negative
trueneg <- spec*isneg
# False Positive
falsepos <- (1-spec)*isneg
# False Negative
falseneg <- (1-sens)*ispos
# Total positive tests
testpos <- sens*(ispos) + (1.00 - spec)*(isneg)
# Prob of having HIV given you already had a positive test
probofHIVwithpostest <- truepos/testpos
```
**The theoretical probability of:**   
*A true positive is `r truepos` vs. a false positive is `r falsepos`.*    
    
*A true negative is `r trueneg` vs. a false negative is `r falseneg`.*

```{r}
### Simulation ###
X <- 1000                        # select this many people

# Pick X random people and see if they actually have HIV (1 = yes, 0 = no)
set.seed(124)
yesno <- c(1,0)                  # 1 = yes, has HIV, 0 = no, does not have HIV
HIVprob <- c(0.03,0.97)          # 3 people out of 100 has HIV
subject <- sample(yesno, X, replace = TRUE, prob = HIVprob)
subject
subswithHIV <- sum(subject) 

# Run the HIV detection test to see if it is positive
TF <- c(1,0)
testprob <- c(0.98,0.02)         # test shows true result 98% of the time
testresult <- sample(TF, X, replace = TRUE, prob = testprob)
testresult
truetest <- sum(testresult)

# Count number of times test is pos AND the person has HIV
totpos <- 0
actpos <- 0
for (i in 1:X){
  if (subject[i] == 1 & testresult[i] == 1) # subject has HIV and test result is accurate
    {totpos <- totpos + 1
    actpos <- actpos + 1}
  else 
    {if (subject[i] == 0 & testresult[i] == 0) # subject is neg but test result is false (false pos)
      {totpos <- totpos + 1}
    }
}
pHIVwPtest <- actpos/totpos
```

**Out of `r X` trials, `r subswithHIV` people had HIV. The test gave the correct result `r truetest` times.**     
*The HIV test was positive `r totpos` times.*        
*The HIV test was positive AND the person had HIV `r actpos` times.*    

**The probability of having HIV when you have a positive test result is: `r pHIVwPtest`.**

```{r}
sessionInfo()
```

