---
title: 'Shonkwiler Problem #6'
author: "Stephanie K Moore"
date: "1/26/2016"
output: html_document
---
## Generating Words from Characters, How Long? ##

6. Generate a sequence of characters from the alphabet {a,b,c,d} with each new character chosen with equal probability and independently of the previous choices.  Now slide a "window" of length 4 along this sequence and record the words that you get.  

How many characters do you expect to have to generate before you see all possible 4^4 = 256 different words?     

Use a simulation to answer this question.  Run the simulation for alphabet sizes of 5, 6, and 7 with window sizes 4 and 5.   
*So to find all possible words, take the alphabet size and raise it to the wordlength: 4^4 = 256, 5^4=625, 6^4=1296, 7^4=2401. If the wordlength is 5, then 4^5=1024, 5^5=3125, 6^5=7776, and 7^5=16807*

**Theoretically you might NEVER get all the possible combinations**     

```{r}
set.seed <- 256
abc <- letters[1:7]
wdlistlen <- numeric()
L <- 1
for (W in 4:5){      # W is window size and word length
  for (A in 4:7){ 
    abet <- abc[1:A]
    n <- A^W   # number of possible "words"
    letseqlen <- (10*n)+(W-1) #to get last few words
    letseq <- vector()
    wordlist <- matrix(data=NA,nrow=n,ncol=1)
    # Create random sequence of letters
    letseq <- paste(sample(abet,letseqlen,replace=TRUE),
                  collapse="")
    # Create a list of words
    for (j in 1:(10*n)){ 
      wordlist[j] <- substr(letseq,j,j+(W-1))}
    srtwds <- sort(wordlist) 
    keepword <- matrix(data=NA,ncol=1) 
    k <- 1
    mstop <- length(srtwds)-1
    # m stops one early since compare is to m+1
    for (m in 1:mstop){ 
      if (k > n) {  # we have found all possible words
        break}
      else {
        if (srtwds[m] != srtwds[m+1]){
          keepword[k] <- srtwds[m]
          k <- k + 1}
      }
    }
    wdlistlen[L] <- length(keepword)
    L <- L + 1 # there will be 8 word lists
  }
}
length(keepword)
wdlistlen
```

```{r}
sessionInfo()
```