---
title: 'Shonkwiler Problem #8'
author: "Stephanie K Moore"
date: "`r Sys.Date()`"
output: html_document
---
## The Drunkard's Walk ##   

8. Use a biased coin to simulate a random walk of 30 steps on a line.  If the coin falls heads (H), take one step to the right, if it falls tails (T), take one step left.  Use Pr(H)=0.6 and Pr(T)=0.4. 

    a) Plot a sample path
    b) Make a histogram for 200 such random walks
    c) Report the sample mean of the 200 walks
    d) Report the sample variance of the 200 walks
    e) Compare the mean and variance to their theoretical values
    
```{r}
n <- 30 # number of steps
RL <- c(-1,1) # -1 = R and +1 = L
probRL <- c(0.6, 0.4)
walk <- sample(RL, n, replace = TRUE, prob = probRL)
walk
path <- cumsum(walk)
plot(path, main = "Walk 30 Steps with P = 0.6 You Go Right\nUp is L and Down is R",type= "b")

# Now simulate 200 such walks and plot a histogram...
walks <- rbinom(200,30,.6)  # 200 trials of 30 steps each
emean <- mean(walks)
evar <- var(walks)
hist(walks, breaks = "Scott", ylab = "# of Walks with X steps Right", xlab = "# of Steps to the Right")

# Theoretical mean and variance:
tmean <- 30*.6
tvar <- n*.6*(.4)   # binom var = n*pi*(1-pi)
```
**The mean final position is `r emean`**     
*The theoretical final position is `r tmean`.*

**The empirical variance is `r evar`**    
*The theoretical variance is `r tvar`.*

**The empirical mean is `r mean(emean)`**    
*The theoretical mean is `r tmean`.*
```{r}
sessionInfo()
```