---
title: "Shonkwiler Problem #1"
author: "Stephanie K Moore"
date: "1/26/2016"
output: html_document
---
## Buffon's Needle ##

1. Perform a large number of trials of Buffon's needle problem and estimate the value of pi.  Plot the error as a function of the number of trials.

**For this problem, I tested both increasing the number of trials while holding the number of needles constant as well as increasing the number of needles and holding the number of trials constant. I wanted to see which one, if any, affected the approximation of pi more and how.**

```{r}
library(ggplot2)
set.seed(222)
# generate x and theta randomly
# set d and L # let L = 1 and d = 2
# find out how many times
d <- 2    # distance between boards
L <- 1    # length of needle
pie <- numeric()   # allocate space for pi values 
terr <- numeric()  # allocate space for error values
###########################################################
#### This section holds the number of needles constant ####
#### at 10,000 and varies the number of trials (t)     ####
###########################################################
n <- 10000                     # number of needles
t <- c(10,100,250,500,1000,2000,4000,6000,8000,10000) 
pie <- numeric()               # allocate space for pi values 
terr <- numeric()
for (i in 1:length(t)){ 
  for (j in 1:t[i]){              
    theta <- (pi/2)*runif(n)   # drop n needles at random angles
    x <- (d/2)*runif(n)        # in random spots on the board
    hits <- (x<L/2*sin(theta)) # count how many cross the line T=1 F=0
    perc <- sum(hits)/n        # find the percent that hit line
    if (perc != 0){
      pie[j] <- 1/perc}        # inverse should approx pi
    else {
      pie[j] == 0} 
  }
  terr[i] <- abs(sum(pie)/t[i] - pi) # amount of error from pi
} 
ggplot(data = as.data.frame(terr), aes(x = t, y = terr)) +
  geom_point(color = "green") +
  theme_bw() +
  labs(x = "# of trials", y = "amount off from pi", title = "Needles = 10,000 With Trials Increasing") 
```


```{r}
########################################################
##### This section holds trials constant at 10,000 #####
##### and varies the number of needles  (n)        #####
########################################################
n <- c(10,100,250,500,1000,2000,4000,6000,8000,10000)
t <- 10000                       # number of trials
for (i in 1:length(n)){ 
  for (j in 1:t){              
    theta <- (pi/2)*runif(n[i])  # drop n needles at random angles
    x <- (d/2)*runif(n[i])       # in random spots on the board
    hits <- (x<L/2*sin(theta))   # cross line? T=1 F=0
    perc <- sum(hits)/n[i]       # find percent that cross line
    if (perc == 0){
      pie[j] <- 0}               # inverse should approx pi
    else {
      pie[j] <- 1/perc} 
  }
  if (is.na(sum(pie))){          # determine the amount of error
    terr[i] <- 0}
  else {
    terr[i] <- abs(sum(pie)/t - pi)}
} 
ggplot(data = as.data.frame(terr), aes(x = n, y = terr)) +
  geom_point(color = "green") +
  theme_bw() +
  labs(x = "# of needles", y = "amount off from pi", 
  title = "Trials = 10,000, Needles Increase\nNeedle Values: 10,100,250,500,1000,2000,4000,6000,8000,10000") 
```

#### Interestingly, increasing the number of needles rather than the number of trials improves the approximation of pi faster and keeps it consistently closer to the actual value than does increasing the number of trials. ####

```{r}
sessionInfo()
```
