---
title: 'Shonkwiler Problem #6'
author: "Stephanie K Moore"
date: "1/26/2016"
output: html_document
---
## Generating Words from Characters, How Long? ##

6. Generate a sequence of characters from the alphabet {a,b,c,d} with each new character chosen with equal probability and independently of the previous choices.  Now slide a "window" of length 4 along this sequence and record the words that you get.  

How many characters do you expect to have to generate before you see all possible 4^4 = 256 different words?     

Use a simulation to answer this question.  Run the simulation for alphabet sizes of 5, 6, and 7 with window sizes 4 and 5.     

**The probability of getting any given combination of 4 letters is (1/4)x(1/4)x(1/4)x(1/4) = 1/256. While you will certainly get A combination of letters, the actual probability of getting a particular combination is zero? Therefore, theoretically you might NEVER get all the possible combinations**     

```{r}
set.seed <- 256
abet <- letters[1:4]
A <- 4
wdlen <- 4     # wd = word length 
n <- A^wdlen   # number of possible "words"
# sequence length must be n+3 to get "n" 4-letter words moving window 1 letter at a time
letseqlen <- (10*n)+3  
letseq <- vector()
wordlist <- matrix(data=NA,nrow=n,ncol=1)
# Create random sequence of letters
letseq <- paste(sample(abet,letseqlen,replace=TRUE),collapse="")
# Create a list of words
for (j in 1:(letseqlen-3)){   
  wordlist[j] <- substr(letseq,j,j+3)
}
srtwds <- sort(wordlist)
seqstop <- length(srtwds)-1  
keepword <- matrix(data=NA,ncol=1)
k <- 1
# m stops one early since compare is to m+1
for (m in 1:seqstop){ 
  if (k > 256) {  # we have found all possible words
    break}
  else {
    if (srtwds[m] != srtwds[m+1]){
      keepword[k] <- srtwds[m]
      k <- k + 1}
  }
}
length(keepword)
```

```{r}
sessionInfo()
```