---
title: "Shonkwiler Homework Problem 4"
author: "Stephanie K Moore"
date: "`r Sys.Date()`"
output: html_document
---
## The Expected Value of a Reward Given By r(x) = 1/x When Rolling a Pair of Dice ##    
#### Comparison of the Theoretical Value With a Simulated Empirical Value ####    

*Suppose you roll a pair of dice and receive a reward equal to 1/roll. Estimate the expected value of the reward by simulation and compare to the theoretical value.*   

So for x = 2 thru 12, the reward (R(x)) is 1/2, 1/3 ... 1/11, 1/12 respectively.   

E(x) = sum for x = 2 thru 12 of  P(X=x) * R(X=x)
so the expected reward = P(2)(1/2) + P(3)(1/3) + ... + P(12)(1/12)

For rolling 2 6-sided die the sample space is 36 and the probabilities for rolling 2 thru 12 are (out of 36):   
1, 2, 3, 4, 5, 6, 5, 4, 3, 2, 1   OR    
P(2,12)=0.02777, P(3,11)=0.05555, P(4,10)=0.083333, P(5,9)=0.11111, P(6,8)=0.13888, P(7)=0.16666
    
```{r}
library(ggplot2)
# Calculate theoretical expected value
tev <- numeric() # expected value of reward for each roll value
tev[1] <- 0 # no such thing as rolling a 1, but it makes the computation easier
for (i in 2:7){
  tev[i] <- (1/i)*(i-1)/36 # ex. E(4) = 1/4 * 3/36 
}
j <- 3
for (i in 8:12){
  tev[i] <- (1/i)*(i-j)/36  # ex. E(9) is 1/9 * (9-5)/36
  j <- j+2
}
tev # This immediately displays the expected value of the reward for each roll ...
thereward <- sum(tev)  # the theoretical expected value of rolling 2 dice
# Now find the empirical expected value...
n <- 500000 # number of trials
die1 <- sample(1:6,n,replace=TRUE)
die2 <- sample(1:6,n,replace=TRUE)
roll <- die1 + die2
countnum <- numeric()
countnum[1] <- 0 # no roll of 1 possible
for (i in 2:12){
   countnum[i] <- length(which(roll == i)) # count up the # of times 2, 3, 4, ... 12 were each rolled
}
countnum[2:12] # This immediately displays how many of 500,000 rolls were 2, 3 ... 11, 12
eev <- numeric() # summing eev will give the empirical expected value of the reward
eev[1] <- 0 # roll can't equal to 1
for (i in 2:12){ 
  if (countnum[i] == 0) {  
    eev[i] <- 0}
  else {
    eev[i] <- (1/i)*countnum[i]/n}
}  
eev[2:12] # This immediately displays the empirical expected value for each roll 2-12 (though 1 is not possible)
empreward <- sum(eev) # ta-da! the empirical expecte value of the reward
```
Dice Roll | 2 |  3 |  4 | 5 | 6 | 7 | 8 | 9 | 10 | 11 | 12
------------------|----|-----|----
Number of Times Rolled | `r countnum[2]` | `r countnum[3]` | `r countnum[4]` | `r countnum[5]` | `r countnum[6]` | `r countnum[7]` | `r countnum[8]` | `r countnum[9]` | `r countnum[10]` | `r countnum[11]` | `r countnum[12]`

**The theoretical expected value of the reward is: `r thereward` dollars.**      
**The empirical expected value of the reward is: `r empreward` dollars.**

```{r}
sessionInfo()
```
