---
title: 'Shonkwiler Problem #8'
author: "Stephanie K Moore"
date: "`r Sys.Date()`"
output: html_document
---
## The Drunkard's Walk ##   

8. Use a biased coin to simulate a random walk of 30 steps on a line.  If the coin falls heads (H), take one step to the right, if it falls tails (T), take one step left.  Use Pr(H)=0.6 and Pr(T)=0.4. 

    a) Plot a sample path
    b) Make a histogram for 200 such random walks
    c) Report the sample mean of the 200 walks
    d) Report the sample variance of the 200 walks
    e) compare the mean and variance to their theoretical values
    
```{r}
n <- 30
step <- vector(length=30)
line <- vector(length=61) 
# Pos 31 is the center (start point) 
pos <-31
for (i in 1:n){
  step[i] <- rbinom(1,1,.6)       #if 1 then H, else T
  if(step[i] == 1){
    pos <- pos + 1
    }
  else{
    pos <- pos - 1
  }
}
step
pos
plot(step,type = "b")
# Now simulate 200 such walks
walks <- rbinom(200,60,.6)
emean <- mean(walks)
evar <- var(walks)
tmean <- 60*.6
tvar <- n*.6*(.4)   # binom var = n*pi*(1-pi)
```
**The mean final position is `r emean`**     
*The theoretical final position is `r tmean`.*

**The mean empirical variance is `r evar`**    
*The theoretical variance is `r tvar`.*

**The mean empirical mean is `r mean(emean)`**    
*The theoretical mean is `r tmean`.*
```{r}
sessionInfo()
```