---
title: 'Shonkwiler Problem #6'
author: "Stephanie K Moore"
date: "`r Sys.Date()`"
output: html_document
---
## Generating Words from Characters, How Long? ##

6. Generate a sequence of characters from the alphabet {a,b,c,d} with each new character chosen with equal probability and independently of the previous choices.  Now slide a "window" of length 4 along this sequence and record the words that you get.  

How many characters do you expect to have to generate before you see all possible 4^4 = 256 different words?     

Use a simulation to answer this question.  Run the simulation for alphabet sizes of 5, 6, and 7 with window (wordlength)  sizes 4 and 5.      

*To find all possible words, take the alphabet size (A) and raise it to the wordlength (W).*    
*$4^4$ = 256, $5^4$ = 625, $6^4$ = 1296, $7^4$ = 2401*    
*$4^5$ = 1024, $5^5$ = 3125, $6^5$ = 7776, $7^5$ = 16807*     

**Theoretically you might NEVER get all the possible combinations**     

```{r}
set.seed(1)
abc <- letters[1:7]
wordspercombo <- numeric()

#####
# This code doesn't count the minimum sequence by single letter needed, rather it creates the sequence length
# as alphabet length raised to word length (+ wordlenght - 1). 
# If this sequence is insufficient, it uses multipliers of the original sequence length.
# IE, if 256 + 3 doesn't work for a 4 letter alphabet making 4-letter words, then it tries 2*256 + 3 then 3*256 + 3 etc
# I considered that to be an acceptable overage for the number of letters needed because I felt that looping to add
# 1 letter to the sequence at a time was too computer intensive, especially since even using multiples, I got 
# numbers that varied with each run (randomization at work!).
#####

nmult <- numeric()            # size of multiplier needed per alpha/word length combo to capture all possible words 
L <- 1
ltrseq <- vector()
maxmult <- 20                 # maximum multiplier we will try before quitting
for (W in 4:5
     ){                       # W is window size (word length)
  for (A in 4:7){             # A is the number of letters in the alphabet
    abet <- abc[1:A]
    n <- A^W                  # number of possible "words"
    for (x in 1:maxmult){
      seqlen <- (x*n)+(W-1) # add W-1 to get last few words
      wordlist <- matrix(data=NA,nrow=x*n,ncol=1)
      # Create a random sequence of letters
      ltrseq <- paste(sample(abet,seqlen,replace=TRUE),
                     collapse="")
      # Move the window along the letter sequence and pick up W letter combinations (words)
      for (j in 1:(x*n)){     # Create x*n words because some will be duplicates
        wordlist[j] <- substr(ltrseq,j,j+(W-1))
      }
      wordlist <- sort(wordlist)  # Sort the list so duplicates show up together
      keepword <- matrix(data=NA,ncol=1)
      k <- 1
      keepword[1] <- wordlist[1]  # 1st word in the list is sure to be unique
      # Need to keep 1st occurrence, not last occurrence in order to get last word in array
      mstop <- length(wordlist)-1
      # m stops one early since compare is to m+1
      for (m in 1:mstop){ 
        if (k > n) {  # we have found all possible words
          break}
        else {
          if (wordlist[m] != wordlist[m+1]){
            keepword[k+1] <- wordlist[m+1]
            k <- k + 1}
        }
      }
      if (length(keepword) == n) {             # all possible words have been found, get out
        wordspercombo[L] <- length(keepword)   # this array should end up with the max words for each list 
        nmult[L] <- x                          # record the multiplier needed for this list
        L <- L + 1                             # bump the index up for the new list
        break
      }  
    }
  }
}
```

**In order to find all 16807 possible unique words for a word length of 5 and a 7-letter alphabet, the sequence generated `r length(wordlist)` words. This means approximately only 1 word out of every `r nmult[8]` words was unique.**    

**This is the 1st 10 and last 10 entries for the last list: (to show it works correctly)**    
*`r keepword[1:10]` and `r keepword[16797:16807]`*    

**The multiplier used (for this randomized trial) for each for all possible 4-letter words with a**   
**4-letter alphabet was `r nmult[1]`**    
**5-letter alphabet was `r nmult[2]`**    
**6-letter alphabet was `r nmult[3]`**    
**7-letter alphabet was `r nmult[4]`**    

**while the multiplier used to find all possible words for 5-letter words with a**   
**4-letter alphabet was `r nmult[5]`**    
**5-letter alphabet was `r nmult[6]`**    
**6-letter alphabet was `r nmult[7]`**    
**7-letter alphabet was `r nmult[8]`**    

```{r}
sessionInfo()
```