---
title: 'Shonkwiler Problem #6'
author: "Stephanie K Moore"
date: "1/26/2016"
output: html_document
---

6. Generate a sequence of characters from the alphabet {a,b,c,d} with each new character chosen with equal probability and independently of the previous choices.  Now slide a "window" of length 4 along this sequence and record the words that you get.  

How many characters do you expect to have to generate before you see all possible 4^4= 256 different words?     

**The probability of getting any given combination of 4 letters is (1/4)x(1/4)x(1/4)x(1/4) = 1/256**    
**Since there are 256 possible words, and you want to get ALL of them, I would expect you would have to generate 256 raised to the 256 IF the window goes 4 letters at a time. If it goes one letter at a time (so that "bcada..." forms both "bcad" and "cada" etc) then ... I'm not sure it changes?**

Use a simulation to answer this question.  Run the simulation for alphabet sizes of 5, 6, and 7 with window sizes 4 and 5.
```{r}
# Generate total possible words in alphabetical order:
abet <- vector("list",4)
abet <- c("a","b","c","d")
wd <- matrix()
i <- 1
for (w in 1:length(abet)){
  for (x in 1:length(abet)){
    for (y in 1:length(abet)){
      for (z in 1:length(abet)){
        wd[i] <- paste(c(abet[w],abet[x],abet[y],abet[z]),collapse="")
        i <- i+1
      }
    }
  }
}
wd
```

```{r}
# Note to self: "letters" is a built-in constant in r made up of the 26 lowercase letters and LETTERS for uppercase
alphabet <- c("a","b","c","d")
n <- 256   # number of possible "words"
# sequence length must= n + 3 to get n 4 letter words out of it moving window 1 letter at a time
wordlen <- 4 # word length
letseq <- vector(length=n+3)
wordlist <- matrix(data=NA,nrow=n,ncol=1)
# Generate "n" number of letters in a sequence
letseq <- paste(sample(alphabet, n+3, replace=TRUE),collapse="")
letseq
# "read" letters 1-4, then 2-5 ... to 256-259 into words
for (j in 1:n){
  wordlist[j] <- substr(letseq,j,j+3)
}
wordlist
```

**So, now how do I compare to see how many unique words I generated?**

```{r}
sessionInfo()
```