---
title: "Shonkwiler Homework Problem 4"
author: "Stephanie K Moore"
date: "Sys.Date()"
output: html_document
---

Problem as stated by Dr. Marland:
4. Suppose a "reward" r(x) is given upon the roll of a pair of dice as follows: r(x) = 1/x, for 2,3, ... , 12.  Estimate the expectation of r by simulation and compare to the theoretical value.

A reward is given for rolling a pair of dice by r(x) = 1/x (where x is the probability of rolling 2, 3, ... 12)  

for x = 2 thru 12   
    r(x) = 36,18,12,9,7.2,6,7.2,9,12,18,36 respectively

Estimate the expectation of r by simulation and compare to the theoretical value.

E(x) = sum of all possible values (value times P(value)) (weighted average)

For rolling 2 6-sided die the sample space is 36 and the probabilities for 2 thru 12 are    
    1/36,2/36,3/36,4/36,5/36,6/36,5/36,4/36,3/36,2/36,1/36
    
So theoretical probabilities are        
    P(2,12)=0.02777, P(3,11)=0.05555, P(4,10)=0.083333, P(5,9)=0.11111, P(6,8)=0.13888, P(7)=0.16666
    
```{r}
library(ggplot2)
n <- 10 # number of trials
for (i in 2:12){
  prb[i-1] <- i*prb[i-1]
}
prb
ev <- sum(prb)  # theoretical expected value = 7

n <- 10      # number of trials
# runif gives uniform probability distrib (normal distrib)
emproll <- sample(2:12,n,replace=TRUE) # empirical simulation
emproll
for (i in 2:12){
  emproll[i-1] <- i*sum(emproll[1:n] == i)/n
}
emproll
empev <- sum(emproll)/n
empev
```

Now for some graphs...

scale_x_discrete(breaks=c("2","3","4","5","6","7","8","9","10","11","12")) +



```{r}
sessionInfo()
```
